### 2025-04-19 base-rates.qmd
20250419 - density plot of rookies vs non-rookies by position

load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

@@

```{r}
rookiesDraftedEarly <- player_stats_seasonal %>% 
  filter(years_of_experience == 0) %>% 
  filter(draftround %in% c(1,2)) %>% 
  arrange(player_id, season) %>% 
  group_by(player_id) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>% 
  arrange(player_display_name)

nonRookiesDraftedEarly <- player_stats_seasonal %>% 
  filter(years_of_experience > 0) %>% 
  filter(draftround %in% c(1,2)) %>% 
  arrange(player_display_name)
```

[EXAMINE FOR ROOKIES WHO WERE DRAFTED IN FIRST TWO ROUNDS OF NFL DRAFT]

[ALSO, FOR A SECONDARY ANALYSIS, SUBSET TO PLAYERS WHO PLAYED A FULL SEASON]

>>>>

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

@@

```{r}
rookiesDraftedEarly <- player_stats_seasonal %>% 
  filter(years_of_experience == 0) %>% 
  filter(draftround %in% c(1,2)) %>% 
  arrange(player_id, season) %>% 
  group_by(player_id) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>% 
  mutate(
    rookie = 0,
    rookieLabel = "Rookie"
  ) %>% 
  arrange(player_display_name)

nonRookiesDraftedEarly <- player_stats_seasonal %>% 
  filter(years_of_experience > 0) %>% 
  filter(draftround %in% c(1,2)) %>% 
  mutate(
    rookie = 1,
    rookieLabel = "Non-rookie"
  ) %>% 
  arrange(player_display_name)

playersDraftedEarly <- bind_rows(
  rookiesDraftedEarly,
  nonRookiesDraftedEarly
) %>% 
  mutate(rookieLabel = factor(rookieLabel, levels = c("Rookie","Non-rookie")))
```

```{r}
#| label: fig-densityPlotRookieVsNonRookieByPosition
#| fig-cap: "Density Plot of Fantasy Points for Rookies Versus Non-Rookies by Position Among Players Who Were Drafted in the First Two Rounds of the National Football League Draft."
#| fig-alt: "Density Plot of Fantasy Points for Rookies Versus Non-Rookies by Position Among Players Who Were Drafted in the First Two Rounds of the National Football League Draft."

ggplot2::ggplot(
  data = playersDraftedEarly %>% filter(position %in% c("QB","RB","WR","TE")),
  aes(
    x = fantasyPoints)) + 
  geom_density(
    aes(
      fill = rookieLabel),
    alpha = 0.5) +
  facet_wrap( ~ position) +
  labs(
    x = "Fantasy Points",
    y = "Density",
    title = "Density Plot of Fantasy Points for Rookies Versus Non-Rookies by Position"
  ) +
  theme_bw() + 
  theme(
    axis.title.y = element_text(angle = 0, vjust = 0.5), # horizontal y-axis title
    legend.title = element_blank()) # remove legend title
```
